---
title: "Drilldown plots for hierarchical data"
output:
  html_document:
    df_print: paged
---

```{r load, message=FALSE, warnings = FALSE}
library(highcharter)
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(zoo)
library(here)
library(purrr)
library(knitr)
```

Drilldown plots can be easily created using `highcharter` which provides an R wrapper around the `highcahrts` javascript library. The structure for the plot is quite simple with the child nested within the parent. [A simple walk through for the wrapper is provided by the the `highcharter` author Joshua Knust](https://jkunst.com/highcharter/articles/drilldown.html). Below I provide an example using the [weekly payroll jobs index](https://www.abs.gov.au/statistics/labour/earnings-and-work-hours/weekly-payroll-jobs-and-wages-australia/latest-release). 

## Setting up the data 

The `readabs` package provides an easy way to download and tidy data from the Australian Bureau of Statistics, including a convenience function for the weekly payroll jobs release. Weekly payroll jobs are are indexed to the week of the 100th confirmed case of COVID-19 in Australia and are detailed at the overall, industry, and subindustry level. These two data sets are shown below (some minor filter has been added to the indsutry data to only give national level data for all people rather than a by state or by sex view)

```{r loadin, message = FALSE, warnings = FALSE}
industry <- here::here("data", "weekly_payroll_industry.csv") %>% 
  read_csv(show_col_types = FALSE) %>% 
  filter(state == "Australia",
         sex == "Persons",
         age == "All ages") %>% 
  select(-series)
  
subindustry <- here::here("data", "weekly_payroll_subindustry.csv") %>% 
  read_csv(show_col_types = FALSE) %>% 
  select(-series) 
```

```{r echo = FALSE}
knitr::kables(list(
  kable(head(filter(industry, industry == "A-Agriculture, forestry & fishing")),
        caption = "Industry level"),
  kable(head(filter(subindustry, industry == "A-Agriculture, forestry & fishing")),
        caption = "subindustry level")
)
)

```

Our primary interest is in understanding the change from one point in time to the current point in time. We will set 3 key comparisons points that we can look at for the visualisation: 

1. Change from the 100th COVID-19 Case
3. Change on the previous month (while the series is currently weekly it is planned to move to a monthly release moving forward so I using change on month and the most immediate comparison)
  
However, to begin with we will just visualise the index for the maximum reference period. To be sure that the same reference period is being used I am first filtering the industry to the max date and then using this maximum date to filter the subindustry data. With the subindustry data we then want to create a nested structure where each child is nested within the parent and named in a way consistent with the parent data. 
```{r}

industry_max <- industry %>% 
  # Filter to maximum date to get most recent reference period
  filter(date == max(date)) 

subindustry_drilldown <- subindustry %>% 
  #Filter to maximum date to get most recent reference period based on the industry dataframe
  filter(date == max(industry_max$date)) %>% 
  #date column isn't needed any more
  select(-date) %>% 
  ## create nested parent child rows 
  group_nest(industry) %>% 
  ## Update teh strcuture so that the toplevel plot and drilldwon speak to eachother
  mutate(
    id = industry,
    type = "column",
    data = map(data, 
               mutate, 
               name = subdivision, 
               y = value),
    data = map(data,
               list_parse)
  )


```

First we can create just the simple bar chart with `highcharter` and the created `industry_max` data 

```{r}
highcharter::hchart(
  industry_max,
  "column",
  hcaes(x = industry, y = value)
)

```

Now we can build on this and create a drilldown into subindustry 
```{r}
highcharter::hchart(
  industry_max, #data for top level is coming from industry max
  "column", #the type of graph is column
   hcaes( #set the aesthetics
    x = industry, #industry along the X axis
    y = value, #value along the y axis
    name = industry, #not actually sure what this does but it is the doco
    drilldown = industry #we will drilldown according the the industry children 
  )
) %>% 
  ## Add in the drilldwon data
   hc_drilldown(
    allowPointDrilldown = TRUE,
    series = list_parse(subindustry_drilldown)
  ) %>% 
  ##Add a title to the plot
  highcharter::hc_title(text = "Index for weekly payroll jobs by industry and industry subdivision") %>% 
  ##And a subtitle giving the date
  highcharter::hc_subtitle(
    text = paste("Week ending", format.Date(max(industry$date),"%d %B %Y"))
  )
```

## Adding two indicators
### Change from 4 weeks ago and change from the 100th case 

Now that we have a simple MVP lets add some more useful information. That is the change from the previous fortnight and change from the 100th case. 

To calculate change on 4 weeks previous we can make use of `dplyr::lag` and set the lag to `4` as the data is weekly. When the publication shifts to monthyl this can be updated to a lag of 1. For the change on the 100th case we know that this is the index value so we can hard code it in at this point. However, when the series  is updated to monthly the index period may change. 


```{r}
industry_change <- industry %>% 
  group_by(industry) %>% 
  mutate(
    month_change = round((value - dplyr::lag(value, 1))/dplyr::lag(value, 1) *100, 2),
    index_change = round((value - 100) / 100 *100, 2)
  ) %>% 
  filter(date == max(industry_max$date)) %>% 
  pivot_longer(
    cols = c(month_change, index_change),
    names_to = "change_from",
    values_to = "pct_change"
  )

subindustry_change <- subindustry %>% 
  group_by(subdivision) %>% 
  mutate(
    month_change = round((value - dplyr::lag(value, 1))/dplyr::lag(value, 1) *100, 2),
    index_change = round((value - 100) / 100 *100, 2)
  ) %>% 
  filter(date == max(industry_max$date)) %>% 
   pivot_longer(
    cols = c(month_change, index_change),
    names_to = "change_from",
    values_to = "pct_change"
  ) %>% 
  #date column isn't needed any more
  select(-date) %>% 
  # remove initial grouping
  ungroup() 

subindustry_index_change <- subindustry_change %>% 
  filter(change_from == "index_change") %>% 
  ## create nested parent child rows 
  group_nest(industry) %>% 
  ## Update teh strcuture so that the toplevel plot and drilldwon speak to eachother
  mutate(
    id = industry,
    type = "column",
    data = map(data, 
               mutate, 
               name = subdivision, 
               y = pct_change,
               ),
    data = map(data,
               list_parse)
  )

```

Here we can view the data for the most recent data 
```{r}

highcharter::hchart(
  filter(industry_change, change_from == "index_change"),
  "column",
   hcaes(
    x = industry,
    y = pct_change,
  )
) %>% 
  highcharter::hc_yAxis(title = "% Change") %>% 
  highcharter::hc_xAxis(title = "Industry")

```

And we can add the drilldown component 
```{r}
change_100 <- highcharter::hchart(
  filter(industry_change, change_from == "index_change"),
  "column", #the type of graph is column
   hcaes( #set the aesthetics
    x = industry, #industry along the X axis
    y = pct_change, #value along the y axis
    name = "industry", #not actually sure what this does but it is the doco
    drilldown = industry #we will drilldown according the the industry children 
  )
) %>% 
  ## Add in the drilldwon data
   hc_drilldown(
    allowPointDrilldown = T,
    series = list_parse(subindustry_index_change)
  ) %>% 
  highcharter::hc_yAxis(title = list(text = "% Change")) %>% 
  highcharter::hc_xAxis(title = list(text = "Industry")) %>% 
  ##Add a title to the plot
  highcharter::hc_title(text = "% Change in STP Jobs from 100th COVID-19 Case" ) %>% 
  ##And a subtitle giving the date
  highcharter::hc_subtitle(
    text = paste("Week ending", format.Date(max(industry$date),"%d %B %Y"))
  )

change_100

htmltools::save_html(change_100, file = here::here("plots", "change_from_100_case.html"))
```

## NOT SURE HOW TO DO A GROUPED DRILLDOWN ##
